---
title: "Storage"
subtitle: "Caches and hashes"
author: "Will Landau"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{storage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r suppression, echo = F}
suppressMessages(suppressWarnings(library(drake)))
clean(destroy = TRUE, verbose = FALSE)
unlink(
  c(
    "Makefile", "report.Rmd", "shell.sh",
    "STDIN.o*", "Thumbs.db",
    "faster_cache", "my_storr"
  ),
  recursive = TRUE
)
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  warning = TRUE
)
```

`Drake`'s `make()` function generates your project's output, and `drake` takes storing this output seriously. This guide explains how `drake` caches and hashes its data, and describes customization options that can increase convenience and speed.

# Caches

When you run `make()`, `drake` stores your imports and targets in a hidden cache.

```{r mtcars_storage}
library(drake)
load_mtcars_example(verbose = FALSE) # Get the code with drake_example("mtcars").
config <- make(my_plan)
```

You can explore your cached data using functions `loadd()`, `readd()`, `cached()`, and others.

```{r explore_mtcars}
head(cached())

head(readd(small))

loadd(large)

head(large)

rm(large) # Does not remove `large` from the cache.
```

By default, these objects live in a hidden `.drake` folder in your working directory. 

```{r getdefaultcache, eval = FALSE}
find_cache()
## [1] "/home/you/project/.drake"

find_project()
## [1] "/home/you/project"
```

`Drake` (via [storr](https://github.com/richfitz/storr)) has an object-like interface to these caches.

```{r oopcache}
cache <- get_cache()

cache$list()

head(cache$get("small"))

tail(cache$get("small", namespace = "meta"))

cache$list_namespaces()
```

Create a new cache of your own with `new_cache()`.

```{r newcache1}
cache2 <- new_cache(path = "cache2")
file.exists("cache2")
```

You can use multiple caches simultaneously, default and non-default alike.

```{r multiplecaches}
config  <- drake_config(cache = cache)
config2 <- drake_config(cache = cache2)
outdated(config)
outdated(config2)
make(my_plan, cache = cache)
make(my_plan, cache = cache2)
```

There are a couple different ways to retrieve caches.

- `get_cache(path = "my_path")` assumes `my_path` is a project root containing a `.drake` folder. If it does not find a `.drake` folder in `my_path`, it searches up through the ancestors of `my_path` until it finds one.
- `this_cache(path = "my_path")` literally assumes `my_path` is the path to the cache, `.drake` folder or not.
- `storr::storr_rds("my_path", mangle_key = TRUE)` is analogous to `this_cache(path = "my_path")`.

```{r usepath}
cache3 <- get_cache(path = getwd()) # Finds the .drake folder in your directory.
head(cache3$list())
cache4 <- this_cache(path = "cache2") # The cache folder is literally called cache2.
head(cache4$list())
```

Destroy caches to remove them from your file system.

```{r rm3}
cache4$destroy()
cache2$list() # Same folder as cache4.
```

See [storr](https://github.com/richfitz/storr) for more on `drake`'s caching infrastructure.

# Hash algorithms

The concept of [hashing](https://en.wikipedia.org/wiki/Hash_function) is central to [storr](https://github.com/richfitz/storr)'s internals. [Storr](https://github.com/richfitz/storr) uses hashes to label stored objects, and `drake` leverages these hashes to figure out which targets are up to date and which ones are outdated. A hash is like a target's fingerprint, so the hash changes when the target changes. Regardless of the target's size, the hash is always the same number of characters.

```{r hashes}
library(digest) # package for hashing objects and files
smaller_data <- 12
larger_data <- rnorm(1000)

digest(smaller_data) # compute the hash

digest(larger_data)
```

However, different hash algorithms vary in length.

```{r compare_algo_lengths}
digest(larger_data, algo = "sha512")

digest(larger_data, algo = "md5")

digest(larger_data, algo = "xxhash64")

digest(larger_data, algo = "murmur32")
```

# Which hash algorithm should you choose?

Hashing is expensive, and unsurprisingly, shorter hashes are usually faster to compute. So why not always use `murmur32`? One reason is the risk of collisions: that is, when two different objects have the same hash. In general, shorter hashes have more frequent collisions. On the other hand, a longer hash is not always the answer. Besides the loss of speed, `drake` and [storr](https://github.com/richfitz/storr) sometimes use hash keys as file names, and long hashes could violate the 260-character cap on Windows file paths. That is why `drake` uses a shorter hash algorithm for internal cache-related file names and a longer hash algorithm for everything else.

```{r justified_hash_choices}
default_short_hash_algo()

default_long_hash_algo()

short_hash(cache)

long_hash(cache)
```

# Select the hash algorithms of the cache

If you want to set the hash algorithms, do so right when the cache is first created.

```{r default_cache_reset}
# cache_path(cache) # Default cache from before. # nolint

# Start from scratch to reset both hash algorithms.
clean(destroy = TRUE)

tmp <- new_cache(
  path = default_cache_path(), # The `.drake/` folder.
  short_hash_algo = "crc32",
  long_hash_algo = "sha1"
)

config <- make(my_plan, verbose = FALSE)

short_hash(config$cache) # xxhash64 is the default_short_hash_algo()

long_hash(config$cache) # sha256 is the default_long_hash_algo()
```

You can change the long hash algorithm without throwing away the cache, but your project will rebuild from scratch. As for the short hash, you are committed until you delete the cache and all its supporting files.

```{r more_cache}
outdated(config) # empty

config$cache <- configure_cache(
  config$cache,
  long_hash_algo = "murmur32",
  overwrite_hash_algos = TRUE
)
```

Below, the targets become outdated because the existing hash keys do not match the new hash algorithm.

```{r newhashmorecache}
config <- drake_config(my_plan, verbose = FALSE, cache = config$cache)
outdated(config)

config <- make(my_plan, verbose = FALSE)

short_hash(config$cache) # same as before

long_hash(config$cache) # different from before
```

# Using `storr` directly

If you want bypass `drake` and generate a cache directly from [storr](https://github.com/richfitz/storr), it is best to do so right from the beginning.

```{r use_storr_directly}
library(storr)
my_storr <- storr_rds("my_storr", mangle_key = TRUE)
new_plan <- drake_plan(simple = sqrt(4))
make(new_plan, cache = my_storr)

cached(cache = my_storr)

readd(simple, cache = my_storr)
```

In addition to `storr_rds()`, `drake` supports in-memory caches created from `storr_environment()`. However, parallel computing is not supported these caches. The `jobs` argument must be 1, and the `parallelism` argument must be either `"mclapply"` or `"parLapply"`. (It is sufficient to leave the default values alone.)

```{r memory_caches}
memory_cache <- storr_environment()
other_plan <- drake_plan(
  some_data = rnorm(50),
  more_data = rpois(75, lambda = 10),
  result = mean(c(some_data, more_data))
)

make(other_plan, cache = memory_cache)

cached(cache = memory_cache)

readd(result, cache = memory_cache)
```

In theory, it should be possible to leverage serious databases using `storr_dbi()`. However, if you use such caches, please heed the following.

1. Be sure you have [storr](https://github.com/richfitz/storr) version 1.1.3 or greater installed.
1. Be careful about parallel computing. For example the `storr::storr_dbi()` cache is not thread-safe. Either use no parallel computing at all or set `parallelism = "future"` with `caching = "master"`. The `"future"` backend is currently experimental, but it allows the master process to do all the caching in order to avoid race conditions.

The following example requires the `DBI` and `RSQLite` packages.

```{r dbi_caches, eval = FALSE}
mydb <- DBI::dbConnect(RSQLite::SQLite(), "my-db.sqlite")
cache <- storr::storr_dbi(
  tbl_data = "data",
  tbl_keys = "keys",
  con = mydb
)
load_mtcars_example() # Get the code with drake_example("mtcars").
unlink(".drake", recursive = TRUE)
make(my_plan, cache = cache)
```

# Cleaning up

If you want to start from scratch, you can `clean()` the cache. Use the `destroy` argument to remove it completely. `cache$del()` and `cache$destroy()` are also options, but they leave output file targets dangling. By contrast, `clean(destroy = TRUE)` removes file targets generated by `drake::make()`. `drake_gc()` and `clean(..., garbage_collection = TRUE)` do garbage collection, and `clean(purge = TRUE)` removes all target-level data, not just the final output values.

```{r cleaning_up}
clean(small, large)

cached() # 'small' and 'large' are gone

clean(destroy = TRUE)

clean(destroy = TRUE, cache = my_storr)
```

```{r cleanup_storage, echo = FALSE}
unlink(c("Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"))
```
