---
title: "Time logging"
author: "William Michael Landau"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{timing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r suppression_timing, echo = F}
suppressMessages(suppressWarnings(library(drake)))
clean(destroy = TRUE)
unlink(c("Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"))
```

# Build times

Thanks to [Jasper Clarkberg](https://github.com/dapperjapper), `drake` records how long it takes to build each target.

```{r timing_intro}
library(drake)
load_basic_example()
make(my_plan, jobs = 2, verbose = FALSE) # See also max_useful_jobs(my_plan).
build_times(digits = 8) # From the cache.
build_times(digits = 8, targets_only = TRUE)
```

For `drake` version 4.1.0 and earlier, `build_times()` just measures the elapsed runtime of each command in `my_plan$command`. For later versions, the build times also account for all the internal operations in `drake:::build()`, such as [storage and hashing](https://github.com/wlandau-lilly/drake/blob/master/vignettes/storage.Rmd).

# Predicting runtime

Drake uses these times to predict the runtime of the next `make()`. At the moment, everything is up to date, so a hypothetical next `make()` should be fast. Here, we only factor in the times of the targets (excluding the imports using `targets_only = TRUE`).

```{r predict_runtime}
predict_runtime(
  my_plan,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
```

But you can also predict the elapsed time of a full runthrough scratch (after `clean()`).

```{r predict_runtime_scratch}
predict_runtime(
  my_plan,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
```

Let's change a dependency to make some targets out of date. Now, even though `from_scatch` is `FALSE`, the next `make()` should take some time.

```{r changedep_timing}
reg2 <- function(d){
  d$x3 <- d$x ^ 3
  lm(y ~ x3, data = d)
}
predict_runtime(
  my_plan,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
```

We can also factor in parallelism using the `future_jobs` argument.

```{r future_jobs}
predict_runtime(
  my_plan,
  future_jobs = 1,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
predict_runtime(
  my_plan,
  future_jobs = 2,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
predict_runtime(
  my_plan,
  future_jobs = 4,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
```

# Rate-limiting targets

To predict the next runtime with multiple parallel jobs, drake makes some assumptions.

1. The outdated targets are spread out evenly over the available jobs. (All targets are used if `from_scratch` is `TRUE`)
1. One job gets all the slowest targets (pessimistic scenario).

Then, `drake` simply takes the targets from the slowest job in each parallelizable stage and sums the corresponding elapsed build times. A parallelizable stage is a column in the workflow graph.

```{r faketiminggraph, eval = FALSE}
# Hover, click, drag, zoom, and pan.
plot_graph(my_plan, width = "100%", height = "500px")
```

<iframe
src = "https://cdn.rawgit.com/wlandau-lilly/drake/b0169ed6/images/reg2.html"
width = "100%" height = "600px" allowtransparency="true" 
style="border: none; box-shadow: none">
</iframe>

You can explore the rate-limiting targets.

```{r rate_limiting_targets}
rate_limiting_times(
  my_plan,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
rate_limiting_times(
  my_plan,
  future_jobs = 2,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
rate_limiting_times(
  my_plan,
  future_jobs = 4,
  from_scratch = TRUE,
  digits = 8,
  verbose = FALSE,
  targets_only = TRUE
)
```

# A word of caution

`Drake` only accounts for the targets with logged build times. If some targets have not been timed, `drake` throws a warning and lists the names of the untimed targets.

```{r endofline_timing, echo = F}
clean(destroy = TRUE)
unlink(c("Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"))
```
