---
title: "Results report: package downloads example"
author: "William Michael Landau"
output: html_document
---

```{r setuppackagesreport, echo = FALSE}
suppressPackageStartupMessages(library(drake))
suppressPackageStartupMessages(library(Ecdat))
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 7,
  fig.height = 4
)
```

This small data analysis project explores some trends in R package downloads over time. The datasets are downloaded using the [cranlogs package](https://github.com/metacran/cranlogs).

```{r cranlogsintrovignette}
library(cranlogs)
cran_downloads(packages = "dplyr", when = "last-week")
```

Above, each count is the number of times `dplyr` was downloaded from the RStudio CRAN mirror on the given day. To stay up to date with the latest download statistics, we need to refresh the data frequently. With `drake`, we can bring all our work up to date without restarting everything from scratch. 

See the [example-packages.Rmd vignette](https://github.com/ropensci/drake/blob/master/vignettes/example-packages.Rmd) for the details. If you wish to tinker with this example yourself, you can generate the [underlying code files](https://github.com/ropensci/drake/tree/master/inst/examples/packages) with `drake_example("packages")` and then step through `run.R`. This `report.Rmd` file does not stand on its own. It is meant to be called through `run.R`. In fact, the `run.R` file has instructions (comments) to walk you through the project.

# Results

We look at download trends in the following packages.

```{r reportlibdrakepkgs}
library(drake)
readd(package_list) # see also loadd() # nolint
```

Using the [cranlogs package](https://github.com/metacran/cranlogs), we calculate the mean number of downloads for each package from the RStudio CRAN mirror. We show these averages for the last month,

```{r reportavgrecent}
readd(averages_recent)
```

and from November 1 to December 1 in 2016.

```{r reportavgolder}
readd(averages_older)
```

We visualize the daily downloads from the last month,

```{r reportplotrecent}
readd(plot_recent)
```

and from November 1 to December 1 in 2016.

```{r reportplotolder}
readd(plot_older)
```
